---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r load, echo = TRUE, warning = FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
unzip(zipfile = "activity.zip", exdir = "./data")
activity <- read.csv("./data/activity.csv", header = TRUE)
activity$date <- as.Date(activity$date)
```

## What is mean total number of steps taken per day?

1) histogram of the total number of steps taken each day
```{r total hist, echo = TRUE, warning = FALSE, message = FALSE}
act1 <- activity %>% group_by(date) %>% summarise(total.steps=sum(steps))
ggplot(act1, aes(x=date, y=total.steps)) + geom_histogram(stat="identity", fill="red")
```

2) Mean and Median for of the total number of steps taken per day
```{r median mean, echo = TRUE}
tmp <- act1 %>% summarise(mean.steps = mean(total.steps, na.rm = TRUE), median.steps = median(total.steps, na.rm = TRUE))
tmp
```
## What is the average daily activity pattern?

1) a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days
```{r interval, echo = TRUE}
act2 <- activity %>% group_by(interval) %>% summarise(mean.steps=mean(steps, na.rm=TRUE))
with(act2, plot(interval, mean.steps, type = "l", col = "red"))
```

2) five-minute interval, contains the maximum number of steps
```{r interval2, echo = TRUE}
head(act2 %>% arrange(desc(mean.steps)), 1)
```
## Imputing missing values

1) total number of missing values in the dataset 
```{r imputing, echo = TRUE}
sum(is.na(activity))
```

2) filling in all of the missing values in the dataset by the mean for that day and  Create a new dataset with not missing values
```{r imputing2, echo = TRUE}
act.imp <- activity
act3 <- activity %>% group_by(wday(date)) %>% summarise(mean.steps=mean(steps, na.rm=TRUE))
for(i in 1:7) {
    act.imp[is.na(act.imp$steps) & wday(act.imp$date) == i, 1] <- act3[i,2]
}
act.imp[is.na(act.imp$steps), 1] <- mean(act.imp$steps, na.rm=T)
```

3) histogram of the total number of steps taken each day
```{r imp hist, echo = TRUE, warning = FALSE, message = FALSE}
act.imp1 <- act.imp %>% group_by(date) %>% summarise(total.steps=sum(steps, na.rm=TRUE))
ggplot(act1, aes(x=date, y=total.steps)) + geom_histogram(stat="identity", fill="blue")
```

4) Mean and Median for of the total number of steps taken per day with imputes missing values
```{r median mean imp, echo = TRUE}
tmp2 <- act.imp1 %>% summarise(mean.steps = mean(total.steps, na.rm = TRUE), median.steps = median(total.steps, na.rm = TRUE))
tmp2
```

5) impact of imputing missing data on the estimates of the total daily number of steps,
difference in the mean and median
```{r mean dif, echo = TRUE}
tmp2[1,1] - tmp[1,1]
```
```{r median dif, echo = TRUE}
tmp2[1,2] - tmp[1,2]
```

## Are there differences in activity patterns between weekdays and weekends?

1) Create a new factor variable in the database
```{r factor, echo = TRUE}
act.imp <- act.imp %>% mutate(type_day = as.factor(ifelse(wday(act.imp$date) %in% c(2,3,4,5,6),"weekday","weekend")))
```

2) time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r week, echo = TRUE}
act4 <- act.imp %>% group_by(interval) %>% filter(type_day == "weekday") %>% summarise(mean.steps.weekdays=mean(steps, na.rm=TRUE))
tmp3 <- act.imp %>% group_by(interval) %>% filter(type_day == "weekend") %>% summarise(mean.steps.weekend=mean(steps, na.rm=TRUE))
act4 <- cbind(act4,tmp3[,2])
with(act4, plot(interval, mean.steps.weekdays, type = "l", col = "red"))
with(act4, points(interval, mean.steps.weekend, type = "l", col = "blue"))
legend("topright", legend = c("mean.steps.weekdays", "mean.steps.weekend"), col = c("red", "blue"), cex = 0.8, pch = "_")
```